---
title: "Pset 3"
author: "Abby Beckler"
date: '2024-02-03'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE, root.dir = "/Users/harshitaramakrishnan/Desktop/Metrics/Earnings_and_Height")
```

# **E4.2**
```{r setup_2, echo=FALSE}
# Load necessary libraries
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(readxl))

# Load the dataset
data <- read_excel("Earnings_and_Height.xlsx")
```

# a. What is the median value of height in the sample?
```{r 42_a}
m_height <- median(data$height)
m_height
```

# b. i. Estimate average earnings for workers whose height is at most 67 inches.
```{r 42_bi}
less_67 <- data$earnings[data$height <= 67]
earn_less_67 <- mean(less_67)
earn_less_67
```

# b. ii. Estimate average earnings for workers whose height is greater than 67 inches.
```{r 42_bii}
more_67 <- data$earnings[data$height > 67]
earn_more_67 <- mean(more_67)
earn_more_67
```

# b. iii. On average, do taller workers earn more than shorter workers? How much more? What is a 95% confidence interval for the difference in average earnings?
```{r 42_biii}
earn_diff = earn_more_67 - earn_less_67
earn_diff
# On average, taller workers earn 5499.44 more than shorter workers
t_test_result <- t.test(less_67, more_67, var.equal = TRUE)
sample_mean <- t_test_result$estimate
t_value <- t_test_result$statistic
standard_error <- t_test_result$stderr
se_times_z <- qt(0.975, df = t_test_result$parameter) * standard_error
ci <-c(sample_mean - se_times_z, sample_mean + se_times_z) %>% 
  round(4) 
paste0("95% Confidence Interval: (", ci[1], ",", ci[2], ")")
```

# c. Construct a scatterplot of annual earnings (Earnings) on height (Height).
```{r 42_c}
ggplot(data, aes(x = height, y = earnings)) +
  geom_point() +
  labs(title = "Scatterplot of Earnings on Height", x = "Height", y = "Earnings")
```

# d. i. Run a regression of Earnings on Height. What is the estimated slope?
```{r 42_di}
reg <- lm(earnings ~ height, data = data)
slope <- coef(reg)[2]
slope
```

# d. ii. Use the estimated regression to predict earnings for a worker who is 67 inches tall, for a worker who is 70 inches tall, and for a worker who is 65 inches tall.
```{r 42_dii}
intercept <- coef(reg)[1]
pred_67 <- slope * 67 + intercept
pred_67
pred_70 <- slope * 70 + intercept
pred_70
pred_65 <- slope * 65 + intercept
pred_65
```

# e. i. - iv. Suppose height were measured in centimeters instead of inches.
# i. What is the estimated slope of the regression?
```{r 42_ei}
data$height_cm <- data$height * 2.54
reg_cm <- lm(earnings ~ height_cm, data = data)
slope_cm <- coef(reg_cm)[2]
slope_cm
```
# ii. What is the estimated intercept?
```{r 42_eii}
intercept_cm <- coef(reg_cm)[1]
intercept_cm
```
# iii. What is the R2?
```{r 42_eiii}
r_squared_cm <- summary(reg_cm)$r.squared
r_squared_cm
```
# iv. What is the standard error of the regression?
```{r 42_eiv}
std_error_cm <- summary(reg_cm)$sigma
std_error_cm
```

# f. i. Run a regression of Earnings on Height, using data for female workers only.
```{r 42_fi}
reg_fm <- lm(earnings ~ height, data = filter(data, sex == 1))
slope_fm <- tidy(reg_fm)$estimate[2]
slope_fm
```

# f. ii. Predict earnings for a woman 1 inch taller than the average woman in the sample.
```{r 42_fii}
taller_fm <- predict(reg_fm, newdata = data.frame(height = mean(data$height[data$sex == 1]) + 1))
taller_fm
```

# g. Repeat (f) for male workers.
# g. i. Run a regression of Earnings on Height, using data for male workers only.
```{r 42_gi}
reg_ml <- lm(earnings ~ height, data = filter(data, sex == 0))
slope_ml <- tidy(reg_ml)$estimate[2]
slope_ml
```

# g. ii. Predict earnings for a man 1 inch taller than the average man in the sample.
```{r 42_gii}
taller_ml <- predict(reg_ml, newdata = data.frame(height = mean(data$height[data$sex == 0]) + 1))
taller_ml
```


# h. Do you think that height is uncorrelated with other factors that cause earning? That is, do you think that the regression error term, u; has a conditional mean of 0 given Height (X)? (You will investigate this more in the Earnings and Height exercises in later chapters.)
If height were uncorrelated with other factors, then differences in height would not correspond to other factors causing differences in earnings. I think this would be true. 




